Hemodynamic responses of the Parahippocampal Place Area to spatial cues in a movie and its audio-description

### 2020-07-08 paper/main.tex
revision of results

\todo[inline]{I revised part of motion correction/co-registration but as usually
I am pretty unsure about correctness}
\todo[inline]{Satz zum Downsampling der 7T Daten?}.

% data sources
% https://github.com/psychoinformatics-de/studyforrest-data-aligned/tree/master/code
% https://github.com/psychoinformatics-de/studyforrest-data-templatetransforms}
fMRI time series of those 15 participants that took part in all three
experiments were obtained from
\href{https://github.com/psychoinformatics-de/studyforrest-data-aligned}{GitHubi}.
Data were already corrected for motion and aligned by non-linear warping to a
particpant-specific BOLD template image \citep{sengupta2016extension}.
% exclusion of VP 10
Data of one participant were dropped to due to invalid distortion correction
during scanning of the AO stimulus.

% preprocessing intro
All further analysis steps of the current study were carried out using tools of
FSL v5.0.9 (\href{https://www.fmrib.ox.ac.uk/fsl}{FMRIB's Software
Library}; \citep{smith2004fsl} on a computer cluster running the
\href{https://www.debian.org}{Debian} GNU/Linux operating system. Software
packages were obtained from repositories of
\href{http://neuro.debian.net}{NeuroDebian} \citep{halchenko2012open}.

@@

% AO: events
For the analysis of the AO stimulus, we created regressors correlating with the
occurrence of descriptive nouns spoken by the narrator that were added to the
original annotation of speech \citep{haeusler2020speechanno}.
% AV events
For the AV stimulus, we created regressors correlating with annotated movie cuts
\citep{haeusler2016cutanno}.
% https://github.com/psychoinformatics-de/studyforrest-data-annotations/blob/master/code/researchcut2segments.py}
First, both annotations were split into eight parts corresponding to the eight
stimulus segments.

% AO events
All of the original 19 categories of descriptive nouns except \texttt{++} were
used as events. Some categories that were similar but offered only a small
amount of counts were pooled resulting in eleven new categories (see Table
\ref{tab:ao-events}). Events that fall into one of the two setting-related
category but also into another category were treated as belonging to
setting-related category (\texttt{se\_new} or \texttt{se\_old}).
% null regressors
If a segment did not contain an event of a category, a null regressor was chosen
for that segment.  \todo{wenn ich "null regressor" nehme, darf ich den ganzen
    Run, glaube ich, in der 2nd level Analyse gar nicht benutzen? Hatte, glaube
    ich, Jeanette Mumford in einem ihren Videos mal gesagt}
% modeling as boxcar function
Events were were modeled as boxcar function from onset to offset of each word.
% categories taken from movie cut annotation
Five additional categories were created from the annotation of movie cuts and
modeled as impulse events to capture variance of a possibly confounding change
of the soundscape after cuts, and to create contrasts of cross-modal negative
control.
% nuissance regressors
Lastly, we used continuous bins of information about two auditory features
(left-right difference in volume and root mean square energy) that was averaged
across the length of every movie frame (\unit[40]{ms}) via Python script to
capture variance correlating with low-level perceptual processes.
% Reference to table
An explanation and counts of the eleven speech-related event categories,
the five movie cut-related categories, and the two low-level confounds can be
found in Table \ref{tab:ao-events}.

@@

% AV events
To create regressors for the AV analysis, we categorized the depicted location
in the frame following one of the 869 cuts into five categories
(cf. Table \ref{tab:av-events}): 1) a cut into an unfamiliar setting
(\texttt{vse\_new}), 2) familiar setting (\texttt{vse\_old}), 3) a switch of
locale within a setting (\texttt{vlo\_ch}), 4) a cut into an unfamiliar
perspective (\texttt{vpe\_new}), and 5) familiar perspective
(\texttt{vpe\_old}).
% null regressor
Here again, a null regressor was chosen for segment if it did not an event for
of a category.
% no cut condition: intro
A sixth regressors (\texttt{no\_cut}) serving a control condition and to created
to build negative control contrasts. Frames within movie shots lasting at least
\unit[20]{s} were pseudo-randomly chosen by the following rationale:
% how they were fitted
\texttt{no\_cut} event had to have a minimum distance of \unit[10]{s} to a movie
cut and to another \texttt{no\_cut} events.
% reference to table
A short explanation and counts of these six categories modeled as impulse
events, six additional event categories for low-level auditory and visual
features of the movie (here again continuous bins lasting \unit[40]{ms}), can be found in Table \ref{tab:av-events}.

@@

% modeling of regressors from events
Events of both stimuli were convolved with a double gamma hemodynamic response
function (HRF).
% reference to figure ``correlation of regressors''
The Pearson correlation coefficients of the 13 regressors of the AO stimulus and
12 regressors of the AV stimulus across the time course of all stimulus segments
can be seen in Figure \ref{fig:reg-corr}.\todo{correct colors on horizontal
line; drop control regressors from figure}
% temporal derivatives
Temporal derivatives were included in the design matrix to compensate for
regional differences between modeled and actual HRF \citep{friston1998event}.
% motion parameters
Six motion parameters were also used as additional nuisance regressors.
% high-pass filtering
Finally, designs were filtered with the same high-pass filter (cutoff period of
\unit[150]{s}) as the BOLD time series.

\begin{figure*} \centering
    \includegraphics[width=\linewidth]{figures/regressor-corr} \caption{Pearson
        correlation coefficients of the regressors used in the analysis of the
        audio-description (blue; see Table \ref{tab:ao-events} for a
        description) and audio-visual movie (red; see Table
        \ref{tab:av-events}). Regressors were created by convolving the events
        with FSL's ``Double-Gamma HRF'' as a model of the hemodynamic response
        function, temporally filtered with the same high-pass filter (cut-off
        \unit[150]{s}) as the BOLD time series, and concatenated across runs
    before computing the correlation.}
    \label{fig:reg-corr}
\end{figure*}

@@

\todo[inline]{Why done this way? Why is no. 1 the one?}
% intro
For the AO stimulus, we created eight $t$-contrasts that aimed to isolate the
PPA and five $t$-contrasts for negative control purposes (see Table \ref{tab:ao-contrasts}).
% PPA contrasts
\todo[inline]{setting-related regressors have small event count compared to
geo/groom; probably more important in discussion}
Contrasts 6-8: words belonging to the categories \texttt{se\_new} and
\texttt{se\_old} were very vague in describing a scene compared to the richness
of a picture. Hence, we assumed smaller probability of successfully isolating
the PPA.
% furniture: blurred border between groom and object
Furniture was not included in one of the contrast because it can be considered
as an object or geometry-defining structure of a room.

% negative control contrasts
Contrasts 9-13 correspond to the contrasts 1-5 of the AV stimulus and served as
negative controls because increased activation in the PPA could also be
correlated with moment of cuts and a possible change in the soundscape of a
scene.

@@

\todo[inline]{change stability of contrast plot from 6 to 5 COPES (line 40 in
script)}

\todo[inline]{Why done this way? Why is no. 1 the one?}
% intro
For the AV stimulus, we created five $t$-contrasts presumably correlating with
increased hemodynamic activity in the PPA by contrasting different kind of cuts.
and five $t$-contrasts for negative control purposes.
% no control for frames content
Notably, this was done without an annotation of the visual content depicted in
the frame after each cut and statistically controlling confounding variables.
% cinematographic rationale
The cinematographic rationale was that movie tend to establishing the setting
(\texttt{vse\_new}) and the spatial relationships within a setting
(\texttt{vpe\_new}) at the beginning of a movie. Later in the movie, the field
sizes of movie shots tend to decrease when switching back to already established
settings (\texttt{vse\_old}). Shots within a recurring settings tend to shift to
depicting persons and objects (\texttt{vpe\_old}) more relevant to the evolved
narrative \citep{brown2012cinematography, mercado2011filmmakers}.
% neuroscientific rationale
The neuroscientific rationale was that simply that activity in the PPA is
greater when participants view novel versus repeated scenes or view-points.
\citep{epstein1999parahippocampal}.
% im Ergebnis
Hence, we created varying contrasts comparing a setting occurring for the first
time and cuts within a locale depicted for the first time to cuts into recurring
settings and cuts within a recurring locale.

@@

\paragraph{alignment etc.}
% single subjects (for Bland-Altman-Plots) 1st level
The GLM analysis that fitted each voxel's time course separately for each
subject was performed in functional space preserving the orientation of the EPI
images.
% two ways of co-registration
For the subsequent second level results, we co-registered the subjects in two
ways.
% subject template (for Bland-Altman-Plot of unthresholded maps)
% \href{"https://github.com/psychoinformatics-de/studyforrest-data-templatetransforms/blob/master/sub-01/bold3Tp2/brain.nii.gz"}{example})
On the one hand, subject-specific BOLD T1* time series were anatomically aligned
via linear transformation to a subject-specific, study-specific T1* template
that served as standard space.
% group template (for group and individual brain slices)
% https://github.com/psychoinformatics-de/studyforrest-data-templatetransforms/blob/master/templates/grpbold3Tp2/brain.nii.gz"
On the other hand, time series were aligned via non-linear transformation to a
study-specific T2* group template.
% MNI152
This reference images was later co-registered to the MNI 152 template with an
affine transformation (12 degrees of freedom) to facilitate visualization and
identification of cortical structures,

@@

% third level
% model
The third level analysis which averaged contrast estimates over subjects was
carried out using FLAME stage 1 with automatic outlier detection
\citep{beckmann2003general, woolrich2004multilevel, woolrich2008robust}.
% thresholding
Here again, Z (Gaussianised T/F) statistic images were thresholded using
clusters determined by Z>3.4 and a (corrected) cluster significance threshold of
p=.05 \citep{worsley2001statistical}.
% brain region identification
Brain regions associated with observed clusters were determined with the Juelich
Histological Atlas \citep{eickhoff2005toolbox, eickhoff2007assignment} and the
Harvard-Oxford Cortical Atlas \citep{desikan2006automated} provided by FSL.
% PPA masks
PPA ROI masks for individual subjects and a PPA group mask of individual PPAs overlapping were created from data provided by \citep{sengupta2016extension}.

@@

% AV
The primary $t$-contrast for the movie comparing cuts to a new setting with cuts
to a familiar perspective (\texttt{vse\_new > vpe\_old}) yielded three
significant clusters (see Table \ref{tab:res-av-group1}, Figure
\ref{fig:group-slices}).
% LOC
Two clusters in the superior lateral occipital cortex (left bigger than right).
% PPA + others.
And one one cluster spanning across the midline and comprising intracalcarine
and cuneal cortex, lingual gyrus and restrosplenial cortex,  occipital fusiform
gyrus and temporal fusiform gyrus, and parahippocampal cortex bilaterally.

@@

>>>>

\todo[inline]{I revised part of motion correction/co-registration but as usually
I am pretty unsure about correctness}
\todo[inline]{Satz zum Downsampling der 7T Daten?}

% data sources
% https://github.com/psychoinformatics-de/studyforrest-data-aligned/tree/master/code
% https://github.com/psychoinformatics-de/studyforrest-data-templatetransforms}
fMRI time series of those 15 participants in the studyforrest dataset that took
part in all three experiments were obtained from
\href{https://github.com/psychoinformatics-de/studyforrest-data-aligned}{GitHub}.
Data were already corrected for motion and aligned by non-linear warping to a
particpant-specific BOLD template image \citep{sengupta2016extension}.
% exclusion of VP 10
Data of one participant were dropped to due to invalid distortion correction
during scanning of the AO stimulus.

% preprocessing intro
All further analysis steps of the current study were carried out using tools of
FSL v5.0.9 (\href{https://www.fmrib.ox.ac.uk/fsl}{FMRIB's Software
Library}; \citep{smith2004fsl}) on a computer cluster running the
\href{https://www.debian.org}{Debian} GNU/Linux operating system. Software
packages were obtained from repositories of
\href{http://neuro.debian.net}{NeuroDebian} \citep{halchenko2012open}.

@@

% AO: events
For the analysis of the AO stimulus, we created regressors correlating with the
occurrence of descriptive nouns spoken by the narrator that were added to the
original annotation of speech \citep{haeusler2020speechanno}.
% AV events
For the AV stimulus, we created regressors correlating with movie cuts provided
by a previously published annotation \citep{haeusler2016cutanno}.
% https://github.com/psychoinformatics-de/studyforrest-data-annotations/blob/master/code/researchcut2segments.py}

% procedure
First, both annotations were split into eight parts corresponding to the eight
stimulus segments.
% AO events
To create regressors for the AV analysis, we used all of the original 19
categories of descriptive nouns except \texttt{++}.
% pooling
Some categories that were similar but offered only a small
amount of counts were pooled resulting in eleven new categories (see Table
\ref{tab:ao-events}). Events that fell into one of the two setting-related
category but also into another category were treated as belonging to the
setting-related category (\texttt{se\_new} or \texttt{se\_old}).
% modeling as boxcar function
Events were were modeled as boxcar function from onset to offset of each word.
% null regressors
If a segment did not contain an event of a category, a null regressor was chosen
for that segment.  \todo{wenn ich "null regressor" nehme, darf ich den ganzen
Run, glaube ich, in der 2nd level Analyse gar nicht benutzen? Hatte, glaube ich,
Jeanette Mumford in einem ihren Videos mal gesagt}
% categories taken from movie cut annotation
Five additional categories were created from the annotation of movie cuts and
modeled as impulse events to capture variance of a possibly confounding change
of the soundscape after a cut, and to create contrasts of cross-modal negative
control.
% nuissance regressors
Lastly, we used continuous bins of information about two auditory features
(left-right difference in volume and root mean square energy) that was averaged
across the length of every movie frame (\unit[40]{ms}) to capture variance
correlating with low-level perceptual processes.
% Reference to table
An explanation and counts of the eleven speech-related event categories, the
five movie cut-related categories, and the two low-level confounds can be found
in Table \ref{tab:ao-events}.

@@

% AV events
To create regressors for the AV analysis, we categorized the depicted location
in the frame following one of the 869 cuts into five categories (see Table
\ref{tab:av-events}): 1) a cut into an unfamiliar setting (\texttt{vse\_new}),
2) a cut into a familiar setting (\texttt{vse\_old}), 3) a switch of locale
within a setting (\texttt{vlo\_ch}), 4) a cut to an unfamiliar perspective
(\texttt{vpe\_new}), and 5) a cut to familiar perspective (\texttt{vpe\_old}).
% null regressor
Here again, a null regressor was chosen for segment if it did not contain an
event of specific category.
% no cut condition: intro
A sixth regressors (\texttt{no\_cut}) serving a control condition and to create
negative control contrasts. Frames within movie shots lasting at least
\unit[20]{s} were pseudo-randomly chosen by the following rationale:
% how they were fitted
\texttt{no\_cut} event had to have a minimum distance of \unit[10]{s} to a movie
cut and to another \texttt{no\_cut} event.
% reference to table
A short explanation and counts of these six categories modeled as impulse
events, and six additional event categories for low-level auditory and visual
features of the movie (here again continuous bins lasting \unit[40]{ms}) can be
found in Table \ref{tab:av-events}.

@@

% modeling of regressors from events
Events of both stimuli were convolved with a double gamma hemodynamic response
function (HRF).
% reference to figure ``correlation of regressors''
The Pearson correlation coefficients of the regressors of the AO and AV stimulus
and across the time course of all stimulus segments can be seen in Figure
\ref{fig:reg-corr}.
% temporal derivatives
Temporal derivatives were included in the design matrix to compensate for
regional differences between modeled and actual HRF \citep{friston1998event}.
% motion parameters
Six motion parameters were also used as additional nuisance regressors.
% high-pass filtering
Finally, designs were filtered with the same high-pass filter (cutoff period of
\unit[150]{s}) as the BOLD time series.

\begin{figure*} \centering
    \includegraphics[width=\linewidth]{figures/regressor-corr} \caption{Pearson
        correlation coefficients of the regressors used in the analysis of the
        audio-description (blue; see Table \ref{tab:ao-events} for a
        description) and audio-visual movie (red; see Table
        \ref{tab:av-events}).
    \label{fig:reg-corr}
\end{figure*}

@@

\todo[inline]{Why done this way? Why is no. 1 the one? Event count,
se\_new and se\_old only vaguely describe the scene; narrator is often
lacking behind dialogue and change in soundscape}

\todo[inline]{contrasts are ordered by ``how well they make sense'' according to
theory, ``quality of auditory cue'', and amount of events/data. Hence, the were
assumed to result in more and more shitty results; at the moment all 8 contrasts
are used in the result (figure)}

% intro
For the AO stimulus, we created eight $t$-contrasts to isolate presumably
correlating with increased hemodynamic activity in the PPA by contrasting
different kind descriptive nouns. % furniture: blurred border between groom and
object Furniture was not included in one of the contrast because it can be
considered as an object or geometry-defining structure of a room.
% negative control contrasts
Five $t$-contrasts (see Table \ref{tab:ao-contrasts} we created for negative
control purposes to test if increased activation in the PPA was correlated with
moment of cuts and therefore correlated with a possible change in the
soundscape.

@@

\todo[inline]{Why done this way? Why is no. 1 the one?}
% intro
For the AV stimulus, we created five $t$-contrasts presumably correlating with
increased hemodynamic activity in the PPA by contrasting different kind of cuts.
% no control for frames content
Notably, this was done without an annotation of the visual content depicted in
the frame after each cut and thus without statistically controlling confounding
variables.
% cinematographic rationale
% cut to new scenes: establishing shots
The cinematographic rationale was that movie tend to establish the setting
(\texttt{vse\_new}) and the spatial relationships within a setting
(\texttt{vpe\_new}) at the beginning of a movie.
% cut to recurrent scene
Later in the movie, the field sizes of movie shots tend to decrease when
switching back to already established settings (\texttt{vse\_old}).
% similarly for shots within a scene
Shots within a recurring settings tend to shift to depicting persons and objects
(\texttt{vpe\_old}) more relevant to the evolved narrative \citep{brown2012cinematography, mercado2011filmmakers}.
% neuroscientific rationale
The neuroscientific rationale was that simply that activity in the PPA is
greater when participants view novel versus repeated scenes or view-points.
\citep{epstein1999parahippocampal}.
% im Ergebnis
Hence, we created varying contrasts comparing a setting occurring for the first
time and cuts within a locale depicted for the first time to cuts into recurring
settings and cuts within a recurring locale.
% negative control
Five additional $t$-contrasts we created for negative control purposes (see
Table \ref{tab:av-contrasts}.

@@

\paragraph{alignment etc.}
% single subjects (for Bland-Altman-Plots) 1st level
The GLM analysis that fitted each voxel's time course separately for each
subject was performed in functional space preserving the orientation of the EPI
images.
% two ways of co-registration
For the subsequent second level results, we co-registered the subjects in two
ways.
% subject template (for Bland-Altman-Plot of unthresholded maps)
% \href{"https://github.com/psychoinformatics-de/studyforrest-data-templatetransforms/blob/master/sub-01/bold3Tp2/brain.nii.gz"}{example})
On the one hand, subject-specific BOLD T1* time series were anatomically aligned
via linear transformation to a subject-specific, study-specific T1* template.
% group template (for group and individual brain slices)
% https://github.com/psychoinformatics-de/studyforrest-data-templatetransforms/blob/master/templates/grpbold3Tp2/brain.nii.gz"
On the other hand, time series were aligned via non-linear transformation to a
study-specific T2* group template.
% MNI152
This reference images was later co-registered to the MNI 152 template with an
affine transformation (12 degrees of freedom) to facilitate visualization and
identification of cortical structures,

@@

% third level
% model
The third level analysis which averaged contrast estimates over subjects was
carried out using FLAME stage 1 with automatic outlier detection
\citep{beckmann2003general, woolrich2004multilevel, woolrich2008robust}.
% thresholding
Here again, Z (Gaussianised T/F) statistic images were thresholded using
clusters determined by Z>3.4 and a (corrected) cluster significance threshold of
p=.05 \citep{worsley2001statistical}.
% brain region identification
Brain regions associated with observed clusters were determined with the Jülich
Histological Atlas \citep{eickhoff2005toolbox, eickhoff2007assignment} and the
Harvard-Oxford Cortical Atlas \citep{desikan2006automated} provided by FSL.
% PPA masks
PPA ROI masks for individual subjects and a PPA group mask of individual PPA
overlaps were created from data provided by \citep{sengupta2016extension}.

@@

% AV
The primary $t$-contrast for the movie comparing cuts to a new setting with cuts
to a familiar perspective (\texttt{vse\_new > vpe\_old}) yielded three
significant clusters (see Table \ref{tab:res-av-group1}, Figure
\ref{fig:group-slices}).
% LOC
Two clusters in the superior lateral occipital cortex (left bigger than right).
% PPA + others.
And one one cluster spanning across the midline and comprising intracalcarine
and cuneal cortex, lingual gyrus and retrosplenial cortex,  occipital fusiform
gyrus and temporal fusiform gyrus, and parahippocampal cortex bilaterally.

@@

Contrasts 6-8: words belonging to the categories \texttt{se\_new} and
\texttt{se\_old} were very vague in describing a scene compared to the richness
of a picture. Hence, we assumed smaller probability of successfully isolating
the PPA.
